---
title: "Lab 3"
author: "Norma Marshall"
format: html

embed-resources: true

fig-width: 10
fig-height: 10


---
```{r}
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)


met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
met <- as.data.frame(met)
```

## **The objectives of the lab are to find the weather station with the highest elevation and look at patterns in the time series of its wind speed and temperature.**

Checking dimensions of dataset 

```{r}
dim(met)
head(met)
tail(met)
```

#**3. Take a look at the variables.
Show the type (class) of each variable (hint: try the str function).
What are the names of the key variables related to our question of interest?**


*Variables of interest are USAFID, elev, year/month/day/hour/min (all measurements of time), wind.sp, and temp.*

```{r}

str(met)
```

**4. Take a closer look at the key variables**

```{r}

table(met$year)
table (met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

It looks like the elevation variable has observations with 9999.0, which is probably an indicator for missing. We should take a deeper look at the data dictionary to confirm. The wind speed variable is OK but there is a lot of missing data.

After checking the data we should make the appropriate modifications. Replace elevations with 9999 as NA.
```{r}
met[met$elev==9999.0, ] <- NA
summary(met$elev)
```

**At what elevation is the highest weather station?**
We also have the issue of the minimum temperature being -40C, which seems implausible, so we should remove those observations.

```{r}
  met <- met[met$temp > -40, ]
  head(met[order(met$temp), ])
```

**6. Calculate summary statistics**
Remember to keep the initial question in mind. We want to pick out the weather station with maximum elevation and examine its wind speed and temperature.


```{r}
elev <- met[which(met$elev == max(met$elev, na.rm = TRUE)), ]
summary(elev)
```


Some ideas: select the weather station with maximum elevation; look at the correlation between temperature and wind speed; look at the correlation between temperature and wind speed with hour and day of the month.

```{r}
cor(elev$temp, elev$wind.sp, use="complete")

cor(elev$temp, elev$hour, use="complete")

cor(elev$wind.sp, elev$day, use="complete")

cor(elev$wind.sp, elev$hour, use="complete")

cor(elev$temp, elev$day, use="complete")
```

##** 7. Exploratory graphs
We should look at the distributions of all of the key variables to make sure there are no remaining issues with the data.**## 

Use the hist function to make histograms of the elevation, temperature, and wind speed variables for the whole dataset

```{r eval=FALSE}

leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)


elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)

elev <- elev[order(elev$date), ]
head(elev)

```

```{r}
hist(elev$temp, 
     main = "Histogram of Elevation", 
     xlab = "Elevation (m)", 
     col = "lightblue", 
     border = "black")
```




